Reading main files and preparing data sets

### 2024-12-10 data-raw/03-preprocessing-main-files.Rmd
Socioeconomic variables from the 2005 GSS survey.

summary(gss_main_2015$DVTDAY)
```
Save data in R format for later use:
```{r}
usethis::use_data(cycling_2015, 
                  overwrite = TRUE)
```

@@

```{r filtering-variables-2010}
gss_main_2010 <- gss_m_2010 %>%
  select(RECID:WGHT_PER, AGEGR5:DVTDAY)
```

@@

summary(gss_main_2010$PRTYPEC)
```
```{r AGECHRYC}

@@

summary(gss_main_2010$DVTDAY)
```
Save data in R format for later use:
```{r}
usethis::use_data(gss_main_2010, 
                  overwrite = TRUE)
```

>>>>

summary(gss_main_2015$DVTDAY)
```

Save data in R format for later use:
```{r}
usethis::use_data(gss_main_2015, 
                  overwrite = TRUE)
```

@@

```{r filtering-variables-2010}
gss_main_2010 <- gss_m_2010 %>%
  select(RECID:DVTDAY)
```

@@

summary(gss_main_2010$PRTYPEC)
```

```{r AGECHRYC}

@@

summary(gss_main_2010$DVTDAY)
```

Save data in R format for later use:
```{r}
usethis::use_data(gss_main_2010, 
                  overwrite = TRUE)
```

## GSS 2005

```{r reading-2010-main-file}
# Main File
gss_m_2005 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2005/gss-12M0019-E-2005-c-19-m_F1.csv"))
```

```{r filtering-variables-2010}
gss_main_2005 <- gss_m_2005 %>%
  select(RECID:DVTDAY)
```

```{r AGEGR5-2005}
values_ordered <- c("15 to 17 years","18 to 19 years","20 to 24 years",
                    "25 to 29 years","30 to 34 years","35 to 39 year","40 to 44 years",
                    "45 to 49 years","50 to 54 years","55 to 59 year","60 to 64 years",
                    "65 to 69 years","70 to 74 years","75 to 79 year","80 years and over")

gss_main_2005 <- gss_main_2005 %>% 
  mutate( # AGEGR5: Age group of respondent (groups of 5)
    AGEGR5 = 
           case_when(AGEGR5 == 1 ~ "15 to 17 years",
                     AGEGR5 == 2 ~ "18 to 19 years",
                     AGEGR5 == 3 ~ "20 to 24 years",
                     AGEGR5 == 4 ~ "25 to 29 years",
                     AGEGR5 == 5 ~ "30 to 34 years",
                     AGEGR5 == 6  ~ "35 to 39 year",
                     AGEGR5 == 7 ~ "40 to 44 years",
                     AGEGR5 == 8 ~ "45 to 49 years",
                     AGEGR5 == 9 ~ "50 to 54 years",
                     AGEGR5 == 10 ~ "55 to 59 year",
                     AGEGR5 == 11 ~ "60 to 64 years",
                     AGEGR5 == 12 ~ "65 to 69 years",
                     AGEGR5 == 13 ~ "70 to 74 years",
                     AGEGR5 == 14 ~ "75 to 79 year",
                     AGEGR5 == 15 ~ "80 years and over"), 
    AGEGR5 = factor(AGEGR5, levels = values_ordered, ordered = TRUE))

summary(gss_main_2005$AGEGR5)
```

```{r AGEGR10-2005}
values_ordered <- c("15 to 24 years","25 to 34 years","35 to 44 years",
                    "45 to 54 years", "55 to 64 years","65 to 74 years","75 years and over")

gss_main_2005 <- gss_main_2005 %>% 
  mutate( # AGEGR10: Age group of respondent (groups of 10)
    AGEGR10 = 
           case_when(AGEGR10 == 1 ~ "15 to 24 years", 
                     AGEGR10 == 2 ~ "25 to 34 years",
                     AGEGR10 == 3 ~ "35 to 44 years",
                     AGEGR10 == 4 ~ "45 to 54 years",
                     AGEGR10 == 5 ~ "55 to 64 years",
                     AGEGR10 == 6  ~ "65 to 74 years",
                     AGEGR10 == 7 ~ "75 years and over"), 
    AGEGR10 = factor(AGEGR10, levels = values_ordered, ordered = TRUE))

summary(gss_main_2005$AGEGR10)
```

```{r SEX-2005}
values_ordered <- c("Male", "Female")

gss_main_2005 <- gss_main_2005 %>% 
  mutate( # SEX: Sex of respondent
    SEX = 
           case_when(SEX == 1 ~ "Male", 
                     SEX == 2 ~ "Female"), 
    SEX = factor(SEX, levels = values_ordered, ordered = TRUE))

summary(gss_main_2005$SEX)
```

```{r MARSTAT}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # MARSTAT: Marital status of the respondent
    MARSTAT = 
           case_when(MARSTAT == 1 ~ "Married", 
                     MARSTAT == 2 ~ "Living common-law",
                     MARSTAT == 3 ~ "Widowed",
                     MARSTAT == 4 ~ "Separated",
                     MARSTAT == 5 ~ "Divorced",
                     MARSTAT == 6  ~ "Single, never married",
                     MARSTAT == 8 ~ "Don't know",
                     MARSTAT == 9 ~ "Not stated"), 
    MARSTAT = factor(MARSTAT))

summary(gss_main_2005$MARSTAT)
```

```{r AGEPRGRDIF-2005}
values_ordered <- c("Respondent is 11 + years older", 
                    "Respondent is 6 - 10 years older",
                    "Respondent is 5 years older",
                    "Respondent is 4 years older",
                    "Respondent is 3 years older",
                    "Respondent is 2 years older",
                    "Respondent is 1 years older",
                    "Respondent and spouse/partner are the same age",
                    "Respondent is 1 years younger",
                    "Respondent is 2 years younger",
                    "Respondent is 3 years younger",
                    "Respondent is 4 years younger",
                    "Respondent is 5 years younger",
                    "Respondent is 6 - 10 years younger",
                    "Respondent is 11 + years younger",
                    "No spouse/partner in household")

gss_main_2005 <- gss_main_2005 %>% 
  mutate( # AGEPRGRDIF: Age difference between respondent and spouse/partner.
    AGEPRGRDIF = 
           case_when(AGEPRGRDIF == 1 ~ "Respondent is 11 + years older", 
                     AGEPRGRDIF == 2 ~ "Respondent is 6 - 10 years older",
                     AGEPRGRDIF == 3 ~ "Respondent is 5 years older",
                     AGEPRGRDIF == 4 ~ "Respondent is 4 years older",
                     AGEPRGRDIF == 5 ~ "Respondent is 3 years older",
                     AGEPRGRDIF == 6  ~ "Respondent is 2 years older",
                     AGEPRGRDIF == 7  ~ "Respondent is 1 years older",
                     AGEPRGRDIF == 8  ~ "Respondent and spouse/partner are the same age",
                     AGEPRGRDIF == 9  ~ "Respondent is 1 years younger",
                     AGEPRGRDIF == 10  ~ "Respondent is 2 years younger",
                     AGEPRGRDIF == 11  ~ "Respondent is 3 years younger",
                     AGEPRGRDIF == 12  ~ "Respondent is 4 years younger",
                     AGEPRGRDIF == 13  ~ "Respondent is 5 years younger",
                     AGEPRGRDIF == 14  ~ "Respondent is 6 - 10 years younger",
                     AGEPRGRDIF == 15  ~ "Respondent is 11 + years younger",
                     AGEPRGRDIF == 97 ~ "No spouse/partner in household"), 
    AGEPRGRDIF = factor(AGEPRGRDIF, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2005$AGEPRGRDIF)
```

```{r SEXPR-2005}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # SEXPR: Sex of respondent's spouse/partner living in the household.
    SEXPR = 
           case_when(SEXPR == 1 ~ "Male spouse/partner in the household", 
                     SEXPR == 2 ~ "Female spouse/partner in the household",
                     SEXPR == 0 ~ "Respondent has no partner in the household"), 
    SEXPR = factor(SEXPR))

summary(gss_main_2005$SEXPR)
```

```{r PRTYPEC-2005}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # PRTYPEC: Type of partner respondent has living in the household.
    PRTYPEC = 
           case_when(PRTYPEC == 1 ~ "Respondent has a married partner in the household", 
                     PRTYPEC == 2 ~ "Respondent has a common-law partner in the household",
                     PRTYPEC == 0 ~ "Respondent has no spouse/partner in the household"), 
    PRTYPEC = factor(PRTYPEC))

summary(gss_main_2005$PRTYPEC)
```

```{r AGECHRYC}

gss_main_2005 <- gss_main_2005 %>% 
  mutate( # AGECHRYC: Age of respondent's youngest child in household
           AGECHRYC = case_when(AGECHRYC == 25 ~ "25 and over",
                     AGECHRYC == 97 ~ "Not asked - no children in the household",
                     TRUE ~ as.character(AGECHRYC)), 
    AGECHRYC = factor(AGECHRYC, levels = values_ordered, ordered = TRUE))

summary(gss_main_2005$AGECHRYC)
```

```{r CHRFLAG-2005}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # CHRFLAG:  single child(ren) of the respondent living in the household.
    CHRFLAG = 
           case_when(CHRFLAG == 1 ~ "Yes", 
                     CHRFLAG == 2 ~ "No"), 
    CHRFLAG = factor(CHRFLAG))

summary(gss_main_2005$CHRFLAG)
```
```{r CHRINHSDC-2005}
values_ordered <- c("None", "One",
                     "Two",
                     "Three",
                     "Four or more children")

gss_main_2005 <- gss_main_2005 %>% 
  mutate( # CHRINHSDC: Number of respondent's children living in the household (any age or marital status).
    CHRINHSDC = 
           case_when(CHRINHSDC == 0 ~ "None",
                     CHRINHSDC == 1 ~ "One",
                     CHRINHSDC == 2 ~ "Two",
                     CHRINHSDC == 3 ~ "Three",
                     CHRINHSDC == 4 ~ "Four or more children"), 
    CHRINHSDC = factor(CHRINHSDC, levels = values_ordered, ordered = TRUE)) 
summary(gss_main_2005$CHRINHSDC)
```

```{r CHR0014C-2005}
values_ordered <- c("None", "One",
                     "Two",
                     "Three or more children")

gss_main_2005 <- gss_main_2005 %>% 
  mutate( # CHR0014C: Number of respondent's children living in the household (any age or marital status).
    CHR0014C = 
           case_when(CHR0014C == 0 ~ "None",
                     CHR0014C == 1 ~ "One",
                     CHR0014C == 2 ~ "Two",
                     CHR0014C == 3 ~ "Three or more children"), 
    CHR0014C = factor(CHR0014C, levels = values_ordered, ordered = TRUE)) 
summary(gss_main_2005$CHR0014C)
```
```{r CHRTIME6-2005}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # CHRTIME6:  Age group of respondent’s child(ren) in household
    CHRTIME6 = 
           case_when(CHRTIME6 == 1 ~ "No child under 19 years of age at home",
                     CHRTIME6 == 2 ~ "All children under 5 years of age",
                     CHRTIME6 == 3 ~ "All children between 5 and 12 years of age",
                     CHRTIME6 == 4 ~ "All children 13 years of age and older",
                     CHRTIME6 == 5 ~ "At least one child under 5 years of age but not all children",
                     CHRTIME6 == 6 ~ "Other"), 
    CHRTIME6 = factor(CHRTIME6))

summary(gss_main_2005$CHRTIME6)
```

```{r PARHSDC}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # PARHSDC: Type of parents the respondent has within the household.
    PARHSDC = 
           case_when(PARHSDC == 1 ~ "Two biological parents",
                     PARHSDC == 2 ~ "Two parents (biological/adoptive/step)",
                     PARHSDC == 3 ~ "Only one parent",
                     PARHSDC == 7 ~ "Not asked - no parents in the household"), 
    PARHSDC = factor(PARHSDC))

summary(gss_main_2005$PARHSDC)
```

```{r LIVARR08}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # LIVARR08: Living arrangement of respondent’s household (8 categories)
    LIVARR08 = 
           case_when(LIVARR08 == 1 ~ "Under 25 years - Not living with parents",
                     LIVARR08 == 2 ~ "Under 25 years - Living with both parents",
                     LIVARR08 == 3 ~ "Under 25 years - Living with mother only",
                     LIVARR08 == 4 ~ "Under 25 years - Living with father only",
                     LIVARR08 == 5 ~ "25 years of age or older - Not living with parents",
                     LIVARR08 == 6 ~ "25 years of age or older - Living with both parents",
                     LIVARR08 == 7 ~ "25 years of age or older - Living with mother only",
                     LIVARR08 == 8 ~ "25 years of age or older - Living with father only"), 
    LIVARR08 = factor(LIVARR08))

summary(gss_main_2005$LIVARR08)
```

```{r LIVARR12-2005} 
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # LIVARR12: Living arrangement of respondent’s household (12 categories)
    LIVARR12 = 
           case_when(LIVARR12 == 1 ~ "Living alone",
                     LIVARR12 == 2 ~ "Living with sponse only",
                     LIVARR12 == 3 ~ "Living with spouse and single child < 25 years",
                     LIVARR12 == 4 ~ "Living with spouse and single child 25 years or older",
                     LIVARR12 == 5 ~ "Living with spouse and non-child(ren)",
                     LIVARR12 == 6 ~ "Living with spouse and other",
                     LIVARR12 == 7 ~ "Living without spouse and single child  < 25 years",
                     LIVARR12 == 8 ~ "Living without spouse and single child 25 years or older",
                     LIVARR12 == 9 ~ "No spouse and non-single child(ren)",
                     LIVARR12 == 10 ~ "Living with two parents",
                     LIVARR12 == 11 ~ "Living with one parent",
                     LIVARR12 == 12 ~ "Other living arrangement"), 
    LIVARR12 = factor(LIVARR12))

summary(gss_main_2005$LIVARR12)
```

```{r HSDSIZEC-2005}
values_ordered <- c("One person household",
                   "Two person household",
                   "Three person household",
                   "Four person household",
                   "Five person household",
                   "Six or more person household")

gss_main_2005 <- gss_main_2005 %>% 
  mutate( # HSDSIZEC: Household size of respondent
    HSDSIZEC = 
           case_when(HSDSIZEC == 1 ~ "One person household",
                     HSDSIZEC == 2 ~ "Two person household",
                     HSDSIZEC == 3 ~ "Three person household",
                     HSDSIZEC == 4 ~ "Four person household",
                     HSDSIZEC == 5 ~ "Five person household",
                     HSDSIZEC == 6 ~ "Six or more person household"), 
    HSDSIZEC = factor(HSDSIZEC, levels = values_ordered, ordered = TRUE))

summary(gss_main_2005$HSDSIZEC)
```

```{r FAMTYPE-2005}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # FAMTYPE: Respondent's type of family structure.
    FAMTYPE = 
           case_when(FAMTYPE == 1 ~ "Couple only", 
                     FAMTYPE == 2 ~ "Intact family",
                     FAMTYPE == 3 ~ "Step family with common child",
                     FAMTYPE == 4 ~ "Step family without a common child",
                     FAMTYPE == 5 ~ "Lone parent family",
                     FAMTYPE == 7 ~ "No spouse/partner or children in household"), 
    FAMTYPE = factor(FAMTYPE))

summary(gss_main_2005$FAMTYPE)
```

```{r MULTIGEN}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # MULTIGEN:  Three generations or more in the respondent’s household
    MULTIGEN = 
           case_when(MULTIGEN == 1 ~ "Yes", 
                     MULTIGEN == 2 ~ "No",
                     MULTIGEN == 6 ~ "Number of generations unknown"), 
    MULTIGEN = factor(MULTIGEN))

summary(gss_main_2005$MULTIGEN)
```

```{r PRV-2005}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # PRV: Province of residence
    PRV = 
      case_when(PRV == 10 ~ "Newfoundland and Labrador",
                     PRV == 11 ~ "Prince Edward Island",
                     PRV == 12 ~ "Nova Scotia",
                     PRV == 13 ~ "New Brunswick",
                     PRV == 24 ~ "Quebec",
                     PRV == 35 ~ "Ontario",
                     PRV == 46 ~ "Manitoba",
                     PRV == 47 ~ "Saskatchewan",
                     PRV == 48 ~ "Alberta",
                     PRV == 59 ~ "British Columbia"),
    PRV = factor(PRV))

summary(gss_main_2005$PRV) 
```
```{r REGION-2005}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # REGION:  Region of residence of the respondent.
    REGION = 
      case_when(REGION == 1 ~ "Atlantic region",
                     REGION == 2 ~ "Quebec",
                     REGION == 3 ~ "Ontario",
                     REGION == 4 ~ "Prairie region",
                     REGION == 5 ~ "British Columbia"),
    REGION = factor(REGION))

summary(gss_main_2005$REGION)
```

```{r LUC_RST}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # LUC_RST:  Population centre indicator
    LUC_RST = 
      case_when(LUC_RST == 1 ~ "CMA/CA",
                     LUC_RST == 2 ~ "non CMA/CA",
                     LUC_RST == 3 ~ "Prince Edward Island"),
    LUC_RST = factor(LUC_RST))

summary(gss_main_2005$LUC_RST)
```

```{r GTU_Q110-2005}
values_ordered <- c("Every day",
                    "A few times a week",
                    "About once a week",
                    "About once a month",
                    "Less than once a month",
                    "Never", 
                    "Don't know")

gss_main_2005 <- gss_main_2005 %>% 
  mutate( # GTU_Q110: How often do you feel rushed? Would you say it is...?
    GTU_Q110 = 
      case_when(GTU_Q110 == 1 ~ "Every day",
                     GTU_Q110 == 2 ~ "A few times a week",
                     GTU_Q110 == 3 ~ "About once a week",
                     GTU_Q110 == 4 ~ "About once a month",
                     GTU_Q110 == 5 ~ "Less than once a month",
                     GTU_Q110 == 6 ~ "Never",
                     GTU_Q110 == 9 ~ "Don't know"),
    GTU_Q110 = factor(GTU_Q110, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2005$GTU_Q110)
```
```{r GTU_Q120-2005}
values_ordered <- c("More rushed",
                    "About the same",
                    "Less rushed",
                    "Don't know")

gss_main_2005 <- gss_main_2005 %>% 
  mutate( # GTU_Q120: How often do you feel you have time on your hands that you don’t know what to do with?
    GTU_Q120 = 
      case_when(GTU_Q120 == 1 ~ "More rushed",
                     GTU_Q120 == 2 ~  "About the same",
                     GTU_Q120 == 3 ~ "Less rushed",
                     GTU_Q120 == 9 ~ "Don't know"),
    GTU_Q120 = factor(GTU_Q120, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2005$GTU_Q120)
```

```{r GTU_130-2005}
values_ordered <- c("Every day",
                    "A few times a week",
                    "About once a week",
                    "About once a month",
                    "Less than once a month",
                    "Never", 
                    "Don't know")

gss_main_2005 <- gss_main_2005 %>% 
  mutate( # GTU_Q130: How often do you feel you have time on your hands that you don’t know what to do with?
    GTU_Q130 = 
      case_when(GTU_Q130 == 1 ~ "Every day",
                     GTU_Q130 == 2 ~ "A few times a week",
                     GTU_Q130 == 3 ~ "About once a week",
                     GTU_Q130 == 4 ~ "About once a month",
                     GTU_Q130 == 5 ~ "Less than once a month",
                     GTU_Q130 == 6 ~ "Never",
                     GTU_Q130 == 9 ~ "Don't know"),
    GTU_Q130 = factor(GTU_Q130, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2005$GTU_Q130)
```
```{r GTU_Q140-2005}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # GTU_Q140:  Do you feel that the days are just too short to do all the things you want?
    GTU_Q140 = 
           case_when(GTU_Q140 == 1 ~ "Yes", 
                     GTU_Q140 == 2 ~ "No",
                     GTU_Q140 == 8 ~ "Not stated", 
                     GTU_Q140 == 9 ~ "Don't know"), 
    GTU_Q140 = factor(GTU_Q140))

summary(gss_main_2005$GTU_Q140)
```

```{r GTU_Q150-2005}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # GTU_Q150: Diary - Reference day - Type
    GTU_Q150 = 
      case_when(GTU_Q150 == 1 ~ "Time with family and friends (spouse, children, boy-friend - girl-friend)",
                     GTU_Q150 == 2 ~ "Relaxation - personal time",
                     GTU_Q150 == 3 ~ "Practising sports",
                     GTU_Q150 == 4 ~ "Crafts or hobbies",
                     GTU_Q150 == 5 ~ "Outdoor activities",
                     GTU_Q150 == 6 ~ "Reading - writing",
                     GTU_Q150 == 7 ~ "Studies",
                     GTU_Q150 == 8 ~ "Work",
                     GTU_Q150 == 9 ~ "Other",
                     GTU_Q150 == 10 ~ "Travel",
                     GTU_Q150 == 11 ~ "Volunteer",
                     GTU_Q150 == 12  ~ "Nothing",
                     GTU_Q150 == 13  ~ "Do household work",
                     GTU_Q150 == 98 ~ "Not stated",
                     GTU_Q150 == 99 ~ "Don't know",
),
    GTU_Q150 = factor(GTU_Q150))

summary(gss_main_2005$GTU_Q150)
```

```{r DVTDAY}
gss_main_2005 <- gss_main_2005 %>% 
  mutate( # DVTDAY: Diary - Reference day - Type
    DVTDAY = 
      case_when(DVTDAY == 1 ~ "Weekday",
                     DVTDAY == 2 ~ "Saturday",
                     DVTDAY == 3 ~ "Sunday"),
    DVTDAY = factor(DVTDAY))

summary(gss_main_2005$DVTDAY)
```
Save data in R format for later use:
```{r}
usethis::use_data(gss_main_2005, 
                  overwrite = TRUE)
```

## GSS 1998
